Exploring the tree of numerical semigroups

### 2013-10-22 Papier/article.tex
Reduction of section 2

\title{Exploring the tree of numerical semigroups}
\author{XXXX}

@@

For a given positive integer $g$, the number of numerical semigroups  is finite and is denoted by $n_g$. 
In  J.A. Sloane's \emph{on-line encyclopedia of integer}~\cite{OEIS} we find the values of $n_g$ for $g\leq 52$. 
These values have been obtain by M. Bras-Amor\'os (view \cite{BrasAmoros2008} for more details for $g\leq 50)$. 
On his home page \cite{Delgado}, M. Delgado  gives the value of $n_{55}$ without specifying the values of $n_{53}$ and $n_{54}$. 
M.~-Bras Amor\'os used a depth first search exploration of the tree of numerical semigroups up to a given genus.
This tree was introduced by J.C. Rosales and al. in \cite{Rosales} and it is the subject of the Section~\ref{S:Tree}.

Starting with all the numerical semigroups of genus $49$ she obtained the number of numerical semigroups of genus $50$ in $18$ days on a pentium D runing at $3$GHz. 
In the package~\tt{NumericalSgs} \cite{NumericalSgps} of \tt{GAP} \cite{GAP}, M. Delgado together with P.A.~Garcia-Sanchez and J. Morais used the same method of exploration.

The paper is divided as follows.
In section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
In Section~\ref{S:DecNumber} we describe a new representation of numerical semigroups that is well suited to the construction of the tree. 
In Section~\ref{S:Algo} we describe an algorithm based on the representation given in Section~\ref{S:DecNumber} and give its complexity. 
Section~\ref{S:Opti} is more technical, and is devoted to the optimisation of the algorithm introduced in Section~\ref{S:Algo}.

\section{The tree of numerical semigroups}
\label{S:Tree}
We first start by some notations.

@@

\begin{defi}
A subset $X$ of a semigroup is a \emph{generating set} of $S$ if every element of $S$ can be express as a sum of elements in $X$. 
\end{defi}

\begin{nota}
A numerical semigroup admitting $X=\{x_1<x_2<...<x_n\}$ as generating set is denoted  by $S=\left<x_1,...,x_\ell\right>$.
\end{nota}

We now introduce a specific generating set.

\begin{defi}
A non-zero element $x$ of a numerical semigroup $S$ is said to be \emph{irreducible} if it cannot be expressed as a sum of two non-zeros elements of $S$.  
We denote by $\Irr(S)$ the set of all irreducible elements of $S$.
\end{defi}

\begin{prop}
Let $S$ be a numerical semigroup.
Then $\Irr(S)$ is the minimal generating set of $S$.
\end{prop}

\begin{proof}
Assume for a contradiction, that there exists an integer  $x$ in $S$ than cannot be decomposed as a sum of irreducible elements. 
We may assume that $x$ is  minimal with this property. 
As  $x$ cannot be irreducible, there exist $y$ and $z$ in~$S\setminus\{0\}$ satisfying $x=y+z$. 
Since we have $y<x$ and $z<x$, the integers $y$ and $z$ can be expressed as a sum of irreducible elements of $S$ and so $x=y+z$ is a sum of irreducible elements, in contradiction to hypothesis.

As irreducible elements of $S$ cannot be decomposed as the sum of two non-zeros integers in $S$, they must occur in each generating set of $S$.
\end{proof}

We recall that Ap\'ery elements of a numerical semigroup $S$ associated to $m(S)$ are the integers $x$ in $S$ such that $x-m(S)$ is no longer in $S$.
We denote by~$\App(S)$ the set of these elements. 
It is well known that the cardinality of $\App(S)$ is exactly $m(S)$ (see~\cite{Book} for example). 
Note that the set $\Irr(S)$ is in included in $\App(S)$.
In particular, the set $\Irr(S)$ is finite and its cardinality is at most $m(S)$.

@@

Let  $S$ be a numerical semigroup. The set $T=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
As  $[c(S)-1,+\infty[$ is included in $T$ we have $c(T)\leq c(S)-1$. 
Therefore every semigroup $S$ of genus $g$ can be obtained from a semigroup $T$ of genus $g-1$ by removing an element of $T$ greater than or equal to $c(T)$.

@@

Conversely, assume that $x$ is irreducible in  $S$. 
As $0$ is never irreducible, the set~$S^x$ contains~$0$. 
Let $a$ and $b$ be two integers belonging to $S^x$. 
The set $S$ is stable under addition, hence $a+b$ lies in $S$. 
As $S$ is equal to $S^x\cup\{x\}$, the integer $a+b$ also lies in $S^x$ except if it is equal to $x$. 
The latter is impossible since $a$ and $b$ are different from $x$ and $x$ is irreducible.
\end{proof}

@@

We construct the tree of numerical semigroups, denoted by $\mathcal{T}$ as follows. 
The root of the tree is the unique semigroup of genus $0$, \ie, $\left<1\right>$ that is equal to $\NN$. 
If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup $S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than $y$.

The above remarks imply that a semigroup $S$ has depth $g$ in $≠mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
We denote by $\mathcal{T}^{\leq g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.

 \begin{figure}[htf!]
 % \[
 % \pstree[,nodesep=3pt]{\sgnode{1}}{
 % 	\pstree[nodesep=3pt]{\sgnode{2,3} \tlput1}{
 % 		\pstree[treesep=4pt,nodesep=2pt]{\sgnode{3,4,5} \tlput2}{
 % 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{4,5,6,7} \tlput3}{
 % 				\pstree{\sgnode{5,6,7,8,9} \tlput4}{}
 % 				\pstree{\sgnode{\gr 4,6,7,9} \tlput5}{}
 % 				\pstree{\sgnode{\gr 4,\gr 5,7} \trput6}{}
 % 				\pstree{\sgnode{\gr 4,\gr 5,\gr 6} \trput7}{}
 % 			}
 % 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{\gr 3,5,7} \tlput4}{
 % 				\pstree{\sgnode{\gr 3,7,8} \tlput 5}{}
 % 				\pstree{\sgnode{\gr 3,\gr 5} \trput 7}{}
 % 			}
 % 			\pstree[treesep=10pt]{\sgnode{\gr 3,\gr 4} \trput5}{\Tn \Tn \Tn}
 % 		}
 % 		\pstree[nodesep=3pt]{~\hspace{-5em}\sgnode{\gr2,5} \trput3}{
 % 			\pstree[nodesep=3pt]{\sgnode{\gr2,7} \trput5}{
 % 				\pstree[nodesep=3pt]{\sgnode{\gr2,9} \trput7}{}
 % 			}
 % 		}
 % 	}
 % }
 % \vspace{-5em}
 % \]

@@

As the reader can check, the main difficulty to characterize the sond of a semigroup is to determine its irreducible elements.
 In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set $\App(S)$ and then the main difficulty is to describe $\App(S_x)$ from $\App(S)$. 
 This approach is elegant but not sufficiently \emph{basic} for our optimisations.

We conclude this section with some basic results on numerical semigroups of a given genus.
Let $S$ be a numerical semigroup. 
We first prove :
\begin{equation}
\label{E:Irr}
x\in \Irr(S)\quad\textrm{implies}\quad x\leq c(S)+m(S).
\end{equation}
If $y$ is a positive integer with $y \geq c(S)+m(S)$ then $y$ lies in $S$ (as $y\geq c(S)$ holds). Moreover, we always have $y-m(S)\geq c(s)$ and so $y$ is not irreducible.

As $\NN\setminus{S}$ contains $\{1,..,m(S)-1\}$ we have 
\begin{equation}
\label{E:m}
m(S)\leq g(S)+1.
\end{equation}
Let $x$ be an element of $S\cap[0,c(S)-1]$. 
The integer $y=c(S)-1-x$ lies in $[0,c(S)-1]$. 
Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
In contradiction with the definition of conductor. 
Thus we define an involution 
\[
\begin{array}{rcl}
\psi:[0,c(S)-1]&\to&[0,c(S)-1]\\
x&\mapsto&c(S)-1-x
\end{array}
\]
that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$.
The cardinality of $[0,c(S)[\setminus S$ is exactly $g(S)$. Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
Since $\psi$ is injective we must have $k\leq g(S)$. 
From the relation 
\[
[0,c(S)[=S\cap[0,c(S)[\sqcup[0,c(S)[\setminus S
\]
 we obtain $c(S)=k+g(S)$ and so 
\begin{equation}
\label{E:c}
c(S)\leq 2 g(S).
\end{equation}

@@

\begin{defi}
Let $S$ be a numerical semigroup.
For every $x$ of $\NN$  we set
\[
D_S(x)=\{y \in S\ |\ x-y\in S\ \text{and}\ 2y\leq x\}
\]
and $d_S(x)=\card D_S(x)$.
We called $d_S(x)$ the $S$-\emph{decomposition number} of $x$.
The application $d_S:\NN\to\NN$ is the $S$-\emph{decomposition numbers function}.
\end{defi}

Assume that $y$ is an element of $D_S(x)$.
By very definition of $D_S(x)$, the integer~$z=x-y$ also belongs to $S$. 
Then $x$ can be decomposed as $x=y+z$ with $y$ and $z$ in~$S$. 
Moreover the condition $2y\leq x$ implies $y\leq z$. 
In other words if we define~$D'_S(x)$ to be the set of all $(y,z)\in S\times S$ with $x=y+z$ and $y\leq z$ then~$D_S(x)$ is the image of $D'_S(x)$ under the projection on the first coordinate.
 Hence~$D_S(x)$ describes how~$x$ can be decomposed as sums of two elements of $S$,-.
 This is why $d_S(x)$ is called the $S$-decomposition number of $x$.

@@

\begin{proof}
By~\eqref{E:c} we have $c(S)\leq g(S)$ and so the $S$-decomposition number of $c(S)$ occurs in $\delta_S$.
Since $c(S)$ is equal to $\max(\NN\setminus S)$, Proposition~\ref{P:delta} implies 
\[
c(S)=1+\max\{i\in[0,...,3G],\ d_S(i)=0\}.
\]
As all elements of $\NN\setminus S$ are smaller than $c(S)$, their $S$-decomposition numbers are in $\delta_S$ and we obtain
\[
g(S)=\card\{i\in[0,..,3G],\ d_S(i)=0\}.
\]
By \eqref{E:m} the relation $m(S)\leq g(S)+1$ holds. 
This implies that the $S$-decomposition number of $m(S)$ appears in $\delta_S$ : 
\[
m(S)=\min\{i\in[0,...,3G],\ d_S(i)>0\}.
\]
Since, by \eqref{E:Irr},  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than $3G$, equations  \eqref{E:m} and \eqref{E:c} give
\[
\Irr(S)=\{i\in[0,...,3G],\ d_S(i)=1\}.
\]
\end{proof}

@@

Let us now prove the complexity statement. 
Since by \eqref{E:c} and \eqref{E:m} we have $x\leq 3G$ together with $m(S)\leq G+1$, each line from \tt{2} to \tt{8} is done in time $O(\log(G))$.
The \textbf{for} loop needs $O(G)$ steps and each step is done in time $O(\log(G))$. 
Summarizing, these results give that the algorithm runs in time $O(\log(G)\times G)$.
\end{proof}

@@

Let us now prove the space complexity statement.
For  this we need to describe the stack through  the run of the algorithm.
Since the stack is filled  with a depth first search algorithm, it has two properties.
The first one is that reading the stack from the bottom to the top, the genus of  semigroup increases.  
The second one is that, for all $g\in[0,G]$, every semigroup of genus $g$ in the stack has the same father.
As the number of sons of a  semigroup $S$ is the number of $S$-irreducible elements in $[c(S),c(S)+m(S)-1]$,  a semigroup $S$ has at most $m(S)$ sons. 
By~\eqref{E:m}, this implies that a semigroup of genus $g$ as at most $g+1$ sons.
Therefore the stack contains at most $g+1$ semigroup of genus $g+1$ for $g\leq G$. 
So the size of the stack is bounded~by 
\[
S=\sum_{g=0}^Gg=\frac{G(G+1)}2
\]
A semigroup is represented by a quadruple $(c(S),g(S),m(S),\delta_S)$. 
By equations~\eqref{E:c} and~\eqref{E:m}, we have $c\leq 2g(S)$ and $m\leq g(S)+1$. 
As $g(S)\leq G$ holds, the integers~$c$, $g$  and $m$ of the representation of $S$ require a memory space in $O(\log(G))$. 
The size of~$\delta_S=(d_S(0),...,d_S(3G))$ is exactly~$3G+1$. 
Each entry of $\delta_S$ is the $S$-decomposition number of an integer smaller than $3G$ and hence
requires $\log(G)$ bytes of memory space.
Therefore the space complexity of $\delta_S$ is in~$O(\log(G)\times G)$, which implies that the space complexity of the \textsc{Count} algorithm is
\[
O(\log(G)\times G\times S)= O(\log(G)\times G^3).
\]
\end{proof}

>>>>

\title{Exploring the tree of numerical semigroups}
\author{Jean Fromentin and  Florent Hivert}

@@

For a given positive integer $g$, the number of numerical semigroups  is finite and is denoted by $n_g$. 
In  J.A. Sloane's \emph{on-line encyclopedia of integer}~\cite{OEIS} we find the values of $n_g$ for $g\leq 52$. 
These values have been obtain by M. Bras-Amor\'os (view \cite{BrasAmoros2008} for more details for $g\leq 50)$. 
On his home page~\cite{Delgado}, M. Delgado  gives the value of $n_{55}$ without specifying the values of $n_{53}$ and~$n_{54}$.

M.~Bras-Amor\'os used a depth first search exploration of the tree of numerical semigroups up to a given genus.
This tree was introduced by J.C. Rosales and al. in \cite{Rosales} and it is the subject of the Section~\ref{S:Tree}.
Starting with all the numerical semigroups of genus $49$ she obtained the number of numerical semigroups of genus $50$ in $18$ days on a pentium D runing at $3$GHz. 
In the package~\tt{NumericalSgs} \cite{NumericalSgps} of \tt{GAP} \cite{GAP}, M. Delgado together with P.A.~Garcia-Sanchez and J. Morais used the same method of exploration.

The paper is divided as follows.
In Section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
In Section~\ref{S:DecNumber} we describe a new representation of numerical semigroups that is well suited to the construction of the tree. 
In Section~\ref{S:Algo} we describe an algorithm based on the representation given in Section~\ref{S:DecNumber} and give its complexity. 
Section~\ref{S:Opti} is more technical, and is devoted to the optimisation of the algorithm introduced in Section~\ref{S:Algo}.

\section{The tree of numerical semigroups}
\label{S:Tree}

We start this section with definitions and properties of numerical semigroups that will be used in the sequel.
For a more complete introduction, the reader can usefully consults the book \emph{Numerical Semigroups} of J.C.~Rosales and P.A.~Garc\'ia-S\'anchez \cite{BookNS} or the book \emph{The Diophantine Frobenius Problem} of J.L.~Ram\'irez~Alfons\'in \cite{BookDFP}.

@@

\begin{defi}
A subset $X=\{x_1<x_2<...<x_n\}$ of a semigroup is a \emph{generating set} of $S$ if every element of $S$ can be express as a sum of elements in $X$. 
In this case we write $S=\left<x_1,...,x_\ell\right>$.
\end{defi}

@@

A non-zero element $x$ of a numerical semigroup $S$ is said to be \emph{irreducible} if it cannot be expressed as a sum of two non-zeros elements of $S$.  
We denote by $\Irr(S)$ the set of all irreducible elements of $S$.

\begin{lem}[Lemma 2.3, Chap. I, \cite{BookNS}]
Let $S$ be a numerical semigroup. The set $\Irr(S)$ is the minimal generating set of $S$.
\end{lem}

%\begin{proof}
%Assume for a contradiction, that there exists an integer  $x$ in $S$ than cannot be decomposed as a sum of irreducible elements. 
%We may assume that $x$ is  minimal with this property. 
%As  $x$ cannot be irreducible, there exist $y$ and $z$ in~$S\setminus\{0\}$ satisfying $x=y+z$. 
%Since we have $y<x$ and $z<x$, the integers $y$ and $z$ can be expressed as a sum of irreducible elements of $S$ and so $x=y+z$ is a sum of irreducible elements, in contradiction to hypothesis. 
%
%As irreducible elements of $S$ cannot be decomposed as the sum of two non-zeros integers in $S$, they must occur in each generating set of $S$.
%\end{proof}

%Let $S$ be a numericla semigroup. As $m(S)$ lies in $S$ the irreducible elements are include in
%\[
%\App(S)=\{x\in S \, |\, x-m(S)\not\in S\}
%\]
%The set $\App(S)$ is the so-called Ap\'ery set relativey to the element $m(S)$.
%
%\begin{lem}[Lemma 2.4, Chap. I, \cite{BookNS}]
%Let $S$ be a numerical semigroup. The Apéry set $\App(S)$ is finite of cardinality $m(S)$.
%\end{lem}
%
%The following result is a direct consequence of the previous Lemma and the Inclusion of $\Irr(S)$ in $m(S)$.
%
%\begin{prop}
%\label{P:Res}
%A numerical semigroup has at most $m(S)$ irreducible elements.
%\end{prop}

The differents parameters we have defined on a numerical semigroup, statisfy the following relations.

\begin{prop}
\label{P:Res}
Let $S$ be a numerical semigroup. We have

$i)$  $x\in \Irr(S)$ implies $x\leq c(S)+m(S)-1$;

$ii)$ $m(S)\leq g(S)+1$;

$iii)$ $c(S)\leq 2 g(S)$.
\end{prop}

\begin{proof}
$i)$ Let $x$ ba a positive integer satisfying $x \geq c(S)+m(S)$. 
As the relations $x\geq c(S)$ and $x-m(S)$ hold, the integers $x$ and $x-m(S)$ belong to $S$.
Since $x=(x-m(S))+m(S)$ hold with $x-m(S)\not=0$ and $m(S)\not=0$, the integer $x$ could not be irreducible.

$ii)$ The set $\NN\setminus{S}$ of cardinality $g(S)$ contains $\{1,..,m(S)-1\}$ and so the relation $m(S)\leq g(S)+1$ holds.

$iii)$ Let $x$ be an element of $S\cap[0,c(S)-1]$. 
The integer $y=c(S)-1-x$ lies in $[0,c(S)-1]$. 
Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
In contradiction with the definition of conductor. 
Thus we define an involution 
\[
\begin{array}{rcl}
\psi:[0,c(S)-1]&\to&[0,c(S)-1]\\
x&\mapsto&c(S)-1-x
\end{array}
\]
that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$.
The cardinality of $[0,c(S)[\setminus S$ is exactly $g(S)$. Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
Since $\psi$ is injective we must have $k\leq g(S)$. 
From the relation 
\[
[0,c(S)[=S\cap[0,c(S)[\sqcup[0,c(S)[\setminus S
\]
 we obtain $c(S)=k+g(S)$ and so $c(S)\leq 2 g(S)$.
\end{proof}

A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite of cardinality at most $c(S)+m(S)-1$.
In fact, using Apéry set, we can prove that  the cardinality of $\Irr(S)$ is at most $m(S)$. See Section 2 of Chapter I of \cite{BookNS} for more details.

Let  $S$ be a numerical semigroup. The set $T=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
As  $[c(S)-1,+\infty[$ is included in $T$ we have $c(T)\leq c(S)-1$. 
Therefore every semigroup $S$ of genus $g$ can be obtained from a semigroup $T$ of genus $g-1$ by removing an element of~$T$ greater than or equal to $c(T)$.

@@

Conversely, assume that $x$ is irreducible in  $S$. 
As $0$ is never irreducible, the set~$S^x$ contains~$0$. 
Let $a$ and $b$ be two integers belonging to $S^x$. 
The set~$S$ is stable under addition, hence $a+b$ lies in $S$. 
As $S$ is equal to $S^x\cup\{x\}$, the integer $a+b$ also lies in $S^x$ except if it is equal to $x$. 
The latter is impossible since $a$ and $b$ are different from $x$ and $x$ is irreducible.
\end{proof}

@@

We construct the tree of numerical semigroups, denoted by $\mathcal{T}$ as follows. 
The root of the tree is the unique semigroup of genus $0$, \ie, $\left<1\right>$ that is equal to $\NN$. 
If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup~$S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than~$y$.

The above remarks imply that a semigroup $S$ has depth $g$ in $\mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
We denote by $\mathcal{T}^{\leq g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.

 \begin{figure}[hf!]
 % \[
 % \pstree[,nodesep=3pt]{\sgnode{1}}{
 % 	\pstree[nodesep=3pt]{\sgnode{2,3} \tlput1}{
 % 		\pstree[treesep=4pt,nodesep=2pt]{\sgnode{3,4,5} \tlput2}{
 % 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{4,5,6,7} \tlput3}{
 % 				\pstree{\sgnode{5,6,7,8,9} \tlput4}{}
 % 				\pstree{\sgnode{\gr 4,6,7,9} \tlput5}{}
 % 				\pstree{\sgnode{\gr 4,\gr 5,7} \trput6}{}
 % 				\pstree{\sgnode{\gr 4,\gr 5,\gr 6} \trput7}{}
 % 			}
 % 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{\gr 3,5,7} \tlput4}{
 % 				\pstree{\sgnode{\gr 3,7,8} \tlput 5}{}
 % 				\pstree{\sgnode{\gr 3,\gr 5} \trput 7}{}
 % 			}
 % 			\pstree[treesep=10pt]{\sgnode{\gr 3,\gr 4} \trput5}{\Tn \Tn \Tn}
 % 		}
 % 		\pstree[nodesep=3pt]{~\hspace{-5em}\sgnode{\gr2,5} \trput3}{
 % 			\pstree[nodesep=3pt]{\sgnode{\gr2,7} \trput5}{
 % 				\pstree[nodesep=3pt]{\sgnode{\gr2,9} \trput7}{}
 % 			}
 % 		}
 % 	}
 % }
 % \vspace{-5em}
 % \]

@@

As the reader can check, the main difficulty to characterize the son of a semigroup is to determine its irreducible elements.
 In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set and then the main difficulty is to describe the Ap\'ery set of $S_x$  from this of $S$. 
 This approach is elegant but not sufficiently \emph{basic} for our optimisations.

@@

\begin{defi}
Let $S$ be a numerical semigroup.
For every $x$ of $\NN$  we set
\[
D_S(x)=\{y \in S\ |\ x-y\in S\ \text{and}\ 2y\leq x\}
\]
and $d_S(x)=\card D_S(x)$.
We called $d_S(x)$ the $S$-\emph{decomposition number} of~$x$.
The application $d_S:\NN\to\NN$ is the $S$-\emph{decomposition numbers function}.
\end{defi}

Assume that $y$ is an element of $D_S(x)$.
By very definition of $D_S(x)$, the integer~$z=x-y$ also belongs to $S$. 
Then $x$ can be decomposed as $x=y+z$ with $y$ and $z$ in~$S$. 
Moreover the condition $2y\leq x$ implies $y\leq z$. 
In other words if we define~$D'_S(x)$ to be the set of all $(y,z)\in S\times S$ with $x=y+z$ and $y\leq z$ then~$D_S(x)$ is the image of $D'_S(x)$ under the projection on the first coordinate.
 Hence~$D_S(x)$ describes how~$x$ can be decomposed as sums of two elements of $S$.
 This is why $d_S(x)$ is called the $S$-decomposition number of $x$.

@@

\begin{proof}
By Proposition~\ref{P:Res} $iii)$ we have $c(S)\leq 2g(S)$ and so the $S$-decomposition number of $c(S)$ occurs in $\delta_S$.
Since $c(S)$ is equal to $\max(\NN\setminus S)$, Proposition~\ref{P:delta} implies 
\[
c(S)=1+\max\{i\in[0,...,3G],\ d_S(i)=0\}.
\]
As all elements of $\NN\setminus S$ are smaller than $c(S)$, their $S$-decomposition numbers are in $\delta_S$ and we obtain
\[
g(S)=\card\{i\in[0,..,3G],\ d_S(i)=0\}.
\]
By Proposition~\ref{P:Res} $ii)$, the relation $m(S)\leq g(S)+1$ holds. 
This implies that the $S$-decomposition number of $m(S)$ appears in $\delta_S$ : 
\[
m(S)=\min\{i\in[0,...,3G],\ d_S(i)>0\}.
\]
Since, by Proposition~\ref{P:Res} $i)$,  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than $3G$ the relations $ii)$ and $iii)$ of Proposition~\ref{P:Res} give
\[
\Irr(S)=\{i\in[0,...,3G],\ d_S(i)=1\}.
\]
\end{proof}

@@

Let us now prove the complexity statement. 
Since by relations $ii)$ and $iii)$ of Proposition~\ref{P:Res} we have $x\leq 3G$ together with $m(S)\leq G+1$, each line from \tt{2} to \tt{8} is done in time $O(\log(G))$.
The \textbf{for} loop needs $O(G)$ steps and each step is done in time $O(\log(G))$. 
Summarizing, these results give that the algorithm runs in time $O(\log(G)\times G)$.
\end{proof}

@@

Let us now prove the space complexity statement.
For  this we need to describe the stack through  the run of the algorithm.
Since the stack is filled  with a depth first search algorithm, it has two properties.
The first one is that reading the stack from the bottom to the top, the genus of  semigroup increases.  
The second one is that, for all $g\in[0,G]$, every semigroup of genus $g$ in the stack has the same father.
As the number of sons of a  semigroup $S$ is the number of $S$-irreducible elements in $[c(S),c(S)+m(S)-1]$,  a semigroup $S$ has at most $m(S)$ sons. 
By~Proposition~\ref{P:Res} $ii)$, this implies that a semigroup of genus $g$ as at most $g+1$ sons.
Therefore the stack contains at most $g+1$ semigroup of genus $g+1$ for $g\leq G$. 
So the size of the stack is bounded~by 
\[
S=\sum_{g=0}^Gg=\frac{G(G+1)}2
\]
A semigroup is represented by a quadruple $(c(S),g(S),m(S),\delta_S)$. 
By relations $ii)$ and $iii)$ of Proposition~\ref{P:Res}, we have $c\leq 2g(S)$ and $m\leq g(S)+1$. 
As $g(S)\leq G$ holds, the integers~$c$, $g$  and $m$ of the representation of $S$ require a memory space in $O(\log(G))$. 
The size of~$\delta_S=(d_S(0),...,d_S(3G))$ is exactly~$3G+1$. 
Each entry of $\delta_S$ is the $S$-decomposition number of an integer smaller than $3G$ and hence
requires $\log(G)$ bytes of memory space.
Therefore the space complexity of $\delta_S$ is in~$O(\log(G)\times G)$, which implies that the space complexity of the \textsc{Count} algorithm is
\[
O(\log(G)\times G\times S)= O(\log(G)\times G^3).
\]
\end{proof}
